---
title: "02Spectra_Data_Cleaning_Exploration"
output: html_document
date: "2023-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### OBJECTIVE of SCRIPT

-   To determine whether the model and the data being regarded (which (generally) won't have associated wet chemistry data, have similar spectral characteristics

    -   So for instance, a soil model could be tested here against some biochar samples to determine whether that soil model could be used to predict biochar properties.

    -   Or, the model we develop for biochar could be overlayed on a new dataset of a biochar (that potentially is of an obscure feedstock) to ensure they overlay in the same spectral space before proceeding.

### Loading Libraries

```{r}
library(prospectr)
library(ggplot2)
library(dplyr)
library(reshape)
library(here)

#setwd("/Users/soils/Library/CloudStorage/OneDrive-CIFOR-ICRAF/Documents/0All_Training/Spectroscopy Data Training-17th May 2023/Data")
```

## Extracting spectra, removing meta data, and briefly visualizing

NOTES

-   Melt functionally takes a dataframe from wide to long format

```{r}

flnm<-"Raw_spectra.csv"

# Read spectra
raw <- read.csv(flnm)
raw[1:2,1:20]
names(raw)[1:20]
dim(raw)

############ Remove spectra metadata ######
raw0 <- raw[,-c(2:17)]
raw0[1:4,1:4]

wavenumbers <- round(as.numeric(substr(colnames(raw0[,-1]),2,100)),1)
##FUSI: this is another way to remove the X that appears in the column names since the values are numeric

colnames(raw0) <- c("SSN", wavenumbers)

raw0[1:5,c(1:4,1700:1704)]

# Create temporary SSNs incase there are repeat scans. The flat csv file retains the original assigned SSNs whether with repeats or not.

length(unique(raw0$SSN))

relabel <- length(grep('TRUE',duplicated(raw0[,1])))

ifelse(relabel>0, raw0[,1] <- paste(raw0$SSN, 1:nrow(raw0),sep = "-"),raw0[,1] <- raw0[,1])

spec.m <- melt(as.data.frame(raw0), id="SSN")


p <- ggplot(data = spec.m, aes(x = as.numeric(as.vector(variable)),y = value,group = SSN)) +
  
  geom_line(size = 0.5, alpha = 0.1, col = 'brown') +
  
  ggtitle(strsplit(flnm,".csv")[[1]][1]) +
  
  xlim(rev(range(as.numeric(as.vector(spec.m$variable))))) +
  
  #note that the variable column refers to the wavenumber and the value column to the absorbance
  ylim(range(spec.m$value)) + 
  
  # ylim(c(0,1.3)) +
  
  xlab(expression("Wavenumbers (cm)"^-1)) +
  
  ylab("Aborbance units") + 
  #theme with white background
  theme_bw() +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
  )
p <- p + theme(plot.title = element_text(hjust = 0.5))

p <- p + theme(legend.position = "none")

p <- p + theme(panel.background = element_rect(fill = "white"))

p

ggsave("Raw_spectra_plot.png")
################################################################################
```

### Cleaning Spectra

NOTES

-   The 'up' and 'down' cleanups are based on:
    -   So thinking about outliers: this is initially scanning the region between 2450-2500 - mainly this is probably a visual scan of the extracted dataset
    -   then maybe one identifies a peak with wavenumber 2498.2 and decides to remove this because it has an absorbance above 1.8
    -   the same is done for the peak at 3696.4 with absorbance below 1.1
    -   then the spectra are recombined

```{r}
################## Clean Spectra ##############

vrbs<-as.numeric(as.vector(spec.m$variable))

##### Up
spec.m[which(vrbs > 2450 & vrbs < 2500),]
bdup<-which(spec.m$variable == 2498.2 & spec.m$value > 1.8)

##### Down
spec.m[which(vrbs > 3600 & vrbs < 3700),]
bddwn<-which(spec.m$variable == 3696.4 & spec.m$value < 1.1)


spec.m2<-spec.m[which(spec.m$SSN %in% spec.m[c(bdup,bddwn),1]),]


bd <- ggplot(data = spec.m2, aes(x = as.numeric(as.vector(variable)),y = value,group = SSN)) +
  
  geom_line(size = 0.5, alpha = 0.1, col = 'brown') + 
  
  xlim(rev(range(as.numeric(as.vector(spec.m2$variable))))) +
  
  ylim(range(spec.m2$value)) + xlab(expression("Wavenumbers (cm)"^-1)) +
  
  ylab("Aborbance units") + theme_bw()

bd

spec.m <- spec.m[which(!spec.m$SSN %in% spec.m[c(bdup,bddwn),1]),]
  
  
############## Average spectra if in duplicate

raw0 <- cast(spec.m, SSN~variable)
dim(raw0)

rownames(raw0) <- raw0$SSN

names(raw0) <- c("SSN", paste0("X", names(raw0)[-1]))

write.csv(raw0, paste0(getwd(),"/","Cleaned_",flnm),row.names = F)
```

```         
```

### Spectral Preprocessing

QUESTIONS

-   The following code seems to be doing the same thing as the 04Overlaying_spectra script - are there main differences?

TO DO

-   Read about the mean-centering function and why it's done here

Notes from Karari -

-   may have to look into database to see how to beef up dataset

-   if biochar does not share similar characteristics to soils data used to build ICRAF models ... ):

```{r}

####################### Read model data and preprocess #########################
master<- read.csv("model-raw_spectra.csv")[,-c(2:17)]
dim(master)
master[1:5,1:5]

wavenumbers <- round(as.numeric(substr(colnames(master[,-1]),2,100)),1)

colnames(master) <- c("SSN", wavenumbers)

#Bind model and new data
nwraw0 <- bind_rows(master,nwraw0)
length(unique(nwraw0$SSN))
################################################################################




nwraw0 <- raw0
##################################### Spectra pre-treatment ###################
#remove the CO2 region from all data
wavenumbers <- colnames(nwraw0[,-1])
co2 <- which(wavenumbers < 2380 & wavenumbers > 2350) # Get co2 bands
nwraw0 <- nwraw0[,-c(co2+1)]
dim(nwraw0)
nwraw0[1:5,1:5]
```

### Savitzky Golay and Mean Centering

```{r}

#SavitzkyGolay pretreatment
spec_der<-as.data.frame(savitzkyGolay(nwraw0[,-1], w=17, p=2, m=1))

#Mean-centering function
center_colmeans <- function(x) {
  xcenter = colMeans(x)
  x - rep(xcenter, rep.int(nrow(x), ncol(x)))
}


#Mean-centering
spec_der.mc_trt<-center_colmeans(spec_der)

spec_trt<-bind_cols(SSN=nwraw0$SSN,spec_der.mc_trt)
 
```

### PCA Plots

TO DO

-   Re-read fundamentals of PCA

```{r}

#Plot PCA scores plot overlay
dsnv <- spec_trt

pcs <- prcomp(dsnv[,-1])

pcss <- pcs$x[,1:10]

pcss[1:6,]

plot(pcss)

calib <- dim(master)[1]

points(pcss[1:calib,1:2], col = "red")

points(pcss[-c(1:calib),1:2], col = "blue")

var <- round(summary(pcs)$importance[2,] * 100, 1)

scores <- cbind("Calib",as.data.frame(pcs$x[,1:5])) # get first 5 principal components

names(scores) <- c("set", colnames(scores[,-1]))

scores <- as.matrix(scores)

scores[-c(1:calib),1] <- "New samples"

scores <- as.data.frame(scores)

write.csv(scores, file = "./Calib and Pred scores.csv", row.names = FALSE)

scores <- read.csv("./Calib and Pred scores.csv")

#sp <- sp +  labs(color = "set")
sp <- ggplot(scores, aes(x = PC1, y =PC2, colour = set)) +
  
  geom_point(size = 0.8, alpha = 0.85 ) +
  
  ggtitle("Calibration PCA scores plot") +
  
  
  xlab(paste0("PC1 explains ", var[1], "% total variance")) +
  
  ylab(paste0("PC2 explains ", var[2], "% total variance")) +
  
  theme_bw() +
  
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
  )
sp <- sp + theme(plot.title = element_text(hjust = 0.5))

sp <- sp + scale_color_manual(values =c("brown","orange"))

sp

ggsave(filename  = "./Soils_scores.png", height = 6, width = 6,sp)
  
```
