---
title: "04Overlaying_spectra"
output: html_document
date: "2023-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overlaying Spectra

QUESTIONS

-   The prepocessing steps are being done over, and done together with the new spectra mainly to ensure they are preprocessed similarly

NOTES

-   Overlaying to check whether model

    -   model = developed using the 100 with wet chem

    -   cleaned_raw is new data without wet chem.cf

-   One hope (and almost assumption) is that soil IR and biochar IR components peak similarly - the way to start validating this is using the 04Overlaying_spectra script

-   before the PLS script- you could be building a model, or building and predicting.

    -   If the former, take out one of the datasets

-   The binding is just to ensure the model and the data being analyzed using the model are preprocessed similarly - they get separated at some point 

```{r}
library(prospectr)
library(ggplot2)
library(dplyr)
library(reshape)

####################### Read model data and preprocess #########################
master<- read.csv("model-raw_spectra.csv")[,-c(2:17)]
master[1:5,1:5]

raw0 <- read.csv("Cleaned_Raw_spectra_training.csv") # this would in actually the new data without wet. hem data 
raw0[1:5,1:5]

#Bind model and new data
nwraw0 <- bind_rows(master,raw0)
length(unique(nwraw0$SSN))
################################################################################

##################################### Spectra pre-treatment ###################
#remove the CO2 region from all data
wavenumbers <- round(as.numeric(substr(colnames(nwraw0[,-1]),2,100)),1)

colnames(nwraw0) <- c("SSN", wavenumbers)

co2 <- which(wavenumbers < 2380 & wavenumbers > 2350) # Get co2 bands
nwraw0 <- nwraw0[,-c(co2+1)]
dim(nwraw0)
nwraw0[1:5,1:5]

#SavitzkyGolay pretreatment
spec_der<-as.data.frame(savitzkyGolay(nwraw0[,-1], w=17, p=2, m=1))

#Mean-centering function
center_colmeans <- function(x) {
  xcenter = colMeans(x)
  x - rep(xcenter, rep.int(nrow(x), ncol(x)))
}


#Mean-centering
spec_der.mc_trt<-center_colmeans(spec_der)

spec_trt<-bind_cols(SSN=nwraw0$SSN,spec_der.mc_trt)


#Plot PCA scores plot overlay
dsnv <- spec_trt

pcs <- prcomp(dsnv[,-1])

pcss <- pcs$x[,1:10]

pcss[1:6,]

plot(pcss)

calib <- dim(master)[1]

points(pcss[1:calib,1:2], col = "red")

points(pcss[-c(1:calib),1:2], col = "blue")

var <- round(summary(pcs)$importance[2,] * 100, 1)

scores <- cbind("Calib",as.data.frame(pcs$x[,1:5])) # get first 5 principal components

names(scores) <- c("set", colnames(scores[,-1]))

scores <- as.matrix(scores)

scores[-c(1:calib),1] <- "New samples"

scores <- as.data.frame(scores)

write.csv(scores, file = "./Calib and Pred scores.csv", row.names = FALSE)

scores <- read.csv("./Calib and Pred scores.csv")

#sp <- sp +  labs(color = "set")
sp <- ggplot(scores, aes(x = PC1, y =PC2, colour = set)) +
  
  geom_point(size = 0.8, alpha = 0.85 ) +
  
  ggtitle("Calibration PCA scores plot") +
  
  
  xlab(paste0("PC1 explains ", var[1], "% total variance")) +
  
  ylab(paste0("PC2 explains ", var[2], "% total variance")) +
  
  theme_bw() +
  
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
  )
sp <- sp + theme(plot.title = element_text(hjust = 0.5))

sp <- sp + scale_color_manual(values =c("brown","orange"))

sp

ggsave(filename  = "./Soils_scores.png", height = 6, width = 6,sp)

```
