## Installing Packages

```{r}

#if (!require("remotes")) install.packages("remotes")
#remotes::install_github("philipp-baumann/simplerspec")

#Info on package: 
  #https://github.com/philipp-baumann/simplerspec/ & 
  #https://github.com/philipp-baumann/simplerspec-read-filter-transform/blob/master/README.Rmd 

#install.packages(c("simplerspec","ggfortify"))
library(devtools)
#install_github("vqv/ggbiplot")
#library(simplerspec)
# Simplerspec is a universal file reader that allows one to read selected parameters instrument, optic and acquisition parameters)

library(readxl)
library(plyr)
library(dplyr)
library(tidyr)
library(ggfortify)
library(tibble)
library(here)
library(ggbiplot)
library(tidyverse)
```

```{r}
files <- list.files("Cornell_MIR_DRIFT_4cm", full.names = TRUE)
str(files)
```

The object `files` has the data structure *atomic vector*. *Atomic vectors* have six possible basic (*atomic*) vector types. These are *logical*, *integer*, *real*, *complex*, *string* (or *character*) and *mir*. Vector types can be returned by the R base function `typeof(x)`, which returns the type or internal storage mode an object `x`. For the `files` object it is

```{r}
# Check type of files object
typeof(files)
```

## Gathering Spectra

-   Took the average of the three scans

-   Created a dataframe (df2) that shows the wavenumber and the average (of 3) absorbances for all samples

```{r}



the_names<- data.frame(union(list.files(path="Cornell_MIR_DRIFT_4cm", pattern="*.csv", full.names=TRUE),list.files(path="Cornell_MIR_DRIFT_4cm", pattern="*.CSV", full.names=TRUE)))
                                                                                    

#naming the column containing file names "ID"
colnames(the_names)=("ID")

#Removing excess characters around the file names to only retain the file name and not the path
#based on the number of strings in the file name (80 in this instance)

#to check the number of characters
nchar(the_names[1,1])

the_names<-the_names%>%
  mutate(ID=substr(ID, 23,32))

absorbance <-union(list.files(path="Cornell_MIR_DRIFT_4cm", pattern="*.csv", full.names=TRUE),list.files(path="Cornell_MIR_DRIFT_4cm", pattern="*.CSV", full.names=TRUE))%>% 
  lapply(read.csv,header=FALSE,col.names=c("wavenumber","absorbance")) %>%
  lapply("[",2) #%>%
 # reduce(merge,by = c("absorbance"))
absorbance

wavenumber <- union(list.files(path="Cornell_MIR_DRIFT_4cm", pattern="*.csv", full.names=TRUE),list.files(path="Cornell_MIR_DRIFT_4cm", pattern="*.CSV", full.names=TRUE)) %>% 
  lapply(read.csv,header=FALSE,col.names=c("wavenumber","absorbance")) %>%
  lapply("[",1) #%>%
 # reduce(merge,by = c("absorbance"))
wavenumber<-wavenumber[[1]]

absorbance<-do.call("cbind", absorbance)
colnames(absorbance)<-the_names[,1]
absorbance

df2<-cbind(wavenumber,absorbance) 
df2<-df2 %>%
  arrange(desc(wavenumber)) 

#uploading the dataframe with the output variables 

#output_vars<-read_excel("IR_Dependent_Variables.xlsx")

```

## Pivoting Dataframe

-   creating wide formatted dataframe (spectra_wide) with ID column (SSN), column with feedstock name (feedstock1), and a column for each wavenumber with the corresponding absorbance as the entry

```{r}
spectra_long<-pivot_longer(df2,c(the_names[,1]))


# The following chubk of code is only valid if you have multiple scans of the same sample, and you want to take the average
# If this is not the case, skip this section
#------------------------------------------------------------------------------------------------------------------------------

newdf <- spectra_long %>%
  group_by(name,wavenumber) %>%
  summarise(value=mean(value,na.rm=FALSE))
newdf<-data.frame(newdf)


wavedf <- spectra_long %>%
  group_by(name) %>%
  summarise(wavenumber=mean(wavenumber,na.rm=FALSE))
wavedf<-data.frame(wavedf)
wavedf<-data.frame(wavedf[,2])
colnames(wavedf)<-("wavenumber")

newdf2<-cbind(wavedf ,newdf)


spectra_wide_2<-pivot_wider(newdf,names_from="wavenumber",values_from="value")
#------------------------------------------------------------------------------------------------------------------------------

spectra_wide<-pivot_wider(spectra_long,names_from="wavenumber",values_from="value")

#renaming identity column SSN to match ICRAF format
colnames(spectra_wide)[1]<-("SSN")
head(spectra_wide)

```

Save file

```{r}
write.csv(spectra_wide, "Raw_spectra_Cornell.csv")
```

## Preliminary PCA

```{r}

#delete_out<-output_vars[c(1:6),]
#delete_in<-data.frame(spectra_wide)

#delete_RHS_pca <-prcomp (delete_in[,c(2:length(delete_in))],
                 #center=TRUE)

tot_pca <- prcomp(spectra_wide[,c(3:length(spectra_wide))],
                 center=TRUE)

#summary(tot_pca)
```

```{r}
g <- ggbiplot::ggbiplot(tot_pca, #need to specify which library the command is coming from, gives errors otherwise
              obs.scale = 1,
              var.scale = 1 ,
              #groups = spectra_wide$feedstock1,
              ellipse = TRUE,
              circle = TRUE,
              ellipse.prob = 0.68,
              var.axes=FALSE)
g
```

# Section 02

---
title: "02Spectra_Data_Cleaning_Exploration"
output: html_document
date: "2023-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### OBJECTIVE of SCRIPT

### Loading Libraries

```{r}
library(prospectr)
library(ggplot2)
library(dplyr)
library(reshape)
library(here)

#setwd("/Users/soils/Library/CloudStorage/OneDrive-CIFOR-ICRAF/Documents/0All_Training/Spectroscopy Data Training-17th May 2023/Data")
```

## Extracting spectra, removing meta data, and briefly visualizing

NOTES

-   Melt functionally takes a dataframe from wide to long format

```{r}

flnm<-"Raw_spectra_Cornell.csv"

# Read spectra
raw <- read.csv(flnm)
raw[1:2,1:20]
names(raw)[1:20]
dim(raw)

############ Remove spectra metadata ######
 
raw0 <- raw[,(2:ncol(raw))] #since the cornell samples have no metadata at present
#raw0 <- raw[,-c(2:17)]
# raw0[1:4,1:4]

wavenumbers <- round(as.numeric(substr(colnames(raw0[,-1]),2,100)),1)
##FUSI: this is another way to remove the X that appears in the column names since the values are numeric

colnames(raw0) <- c("SSN", wavenumbers)

raw0[1:5,c(1:4,1700:1704)]

# Create temporary SSNs incase there are repeat scans. The flat csv file retains the original assigned SSNs whether with repeats or not.

length(unique(raw0$SSN))

relabel <- length(grep('TRUE',duplicated(raw0[,1])))

ifelse(relabel>0, raw0[,1] <- paste(raw0$SSN, 1:nrow(raw0),sep = "-"),raw0[,1] <- raw0[,1])

spec.m <- melt(as.data.frame(raw0), id="SSN")


p <- ggplot(data = spec.m, aes(x = as.numeric(as.vector(variable)),y = value,group = SSN)) +
  
  geom_line(size = 0.5, alpha = 0.1, col = 'brown') +
  
  ggtitle(strsplit(flnm,".csv")[[1]][1]) +
  
  xlim(rev(range(as.numeric(as.vector(spec.m$variable))))) +
  
  #note that the variable column refers to the wavenumber and the value column to the absorbance
  ylim(range(spec.m$value)) + 
  
  # ylim(c(0,1.3)) +
  
  xlab(expression("Wavenumbers (cm)"^-1)) +
  
  ylab("Aborbance units") + 
  #theme with white background
  theme_bw() +
  #eliminates background, gridlines, and chart border
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
  )
p <- p + theme(plot.title = element_text(hjust = 0.5))

p <- p + theme(legend.position = "none")

p <- p + theme(panel.background = element_rect(fill = "white"))

p

ggsave("Raw_spectra_plot_Cornell.png")
################################################################################
```

### Cleaning Spectra

NOTES

-   The 'up' and 'down' cleanups are based on:
    -   So thinking about outliers: this is initially scanning the region between 2450-2500 - mainly this is probably a visual scan of the extracted dataset
    -   then maybe one identifies a peak with wavenumber 2498.2 and decides to remove this because it has an absorbance above 1.8
    -   the same is done for the peak at 3696.4 with absorbance below 1.1
    -   then the spectra are recombined

```{r}
################## Clean Spectra ##############

vrbs<-as.numeric(as.vector(spec.m$variable))

##### Up
spec.m[which(vrbs > 2450 & vrbs < 2500),]
bdup<-which(spec.m$variable == 2498.2 & spec.m$value > 1.8)

##### Down
spec.m[which(vrbs > 3600 & vrbs < 3700),]
bddwn<-which(spec.m$variable == 3696.4 & spec.m$value < 1.1)


#extracting only the rows/samples in the spec.m dataframe that meet the above criteria defined for the bdup and bddwn dataframes
spec.m2<-spec.m[which(spec.m$SSN %in% spec.m[c(bdup,bddwn),1]),]



bd <- ggplot(data = spec.m2, aes(x = as.numeric(as.vector(variable)),y = value,group = SSN)) +
  
  geom_line(size = 0.5, alpha = 0.1, col = 'brown') + 
  
  xlim(rev(range(as.numeric(as.vector(spec.m2$variable))))) +
  
  ylim(range(spec.m2$value)) + xlab(expression("Wavenumbers (cm)"^-1)) +
  
  ylab("Aborbance units") + theme_bw()

bd

#Now redefining the spec.m dataframe to omit the samples that meet the bdup and bddwn criteria as these are defined here as outliers

spec.m <- spec.m[which(!spec.m$SSN %in% spec.m[c(bdup,bddwn),1]),]
  
  
############## Average spectra if in duplicate

raw0 <- cast(spec.m, SSN~variable)
dim(raw0)

rownames(raw0) <- raw0$SSN

names(raw0) <- c("SSN", paste0("X", names(raw0)[-1]))

write.csv(raw0, paste0(getwd(),"/","model_",flnm),row.names = F)
```

```         
```

### Spectral Preprocessing

-   The first chunk is only necessary if you have a model spectra already, and you want to combine it with the new data so you can preprocess it together, you (likely) won't have this in the beginning i.e. if you are working on developing the model so you can skip this chunk

QUESTIONS

-   The following code seems to be doing the same thing as the 04Overlaying_spectra script - are there main differences?

TO DO

-   Read about the mean-centering function and why it's done here

Notes from Karari -

-   may have to look into database to see how to beef up dataset

-   if biochar does not share similar characteristics to soils data used to build ICRAF models ... ):

```{r}
#The following chunk is only necessary if you have a model spectra already, and you want to combine it with the 
# new data so you can preprocess it together, you (likely) won't have this in the beginning i.e. if you are working on developing the model 
# so you can skip this chunk 

####################### Read model data and preprocess #########################
# master<- read.csv("model-raw_spectra.csv")[,-c(2:17)]
# dim(master)
# master[1:5,1:5]
# 
# wavenumbers <- round(as.numeric(substr(colnames(master[,-1]),2,100)),1)
# 
# colnames(master) <- c("SSN", wavenumbers)
# 
# #Bind model and new data
# nwraw0 <- bind_rows(master,nwraw0)
# length(unique(nwraw0$SSN))
################################################################################




nwraw0 <- raw0
##################################### Spectra pre-treatment ###################
#remove the CO2 region from all data
#FUSI EDIT: commenting out this section because need for this dataset, co2 is empty 

# wavenumbers <- colnames(nwraw0[,-1])
# co2 <- which(wavenumbers < 2380 & wavenumbers > 2350) # Get co2 bands
# nwraw0 <- nwraw0[,-c(co2+1)]
# dim(nwraw0)
# nwraw0[1:5,1:5]
```

### Savitzky Golay and Mean Centering

```{r}

#SavitzkyGolay pretreatment
spec_der<-as.data.frame(savitzkyGolay(nwraw0[,-1], w=17, p=2, m=1))

#Mean-centering function
center_colmeans <- function(x) {
  xcenter = colMeans(x)
  x - rep(xcenter, rep.int(nrow(x), ncol(x)))
}


#Mean-centering
spec_der.mc_trt<-center_colmeans(spec_der)

spec_trt<-bind_cols(SSN=nwraw0$SSN,spec_der.mc_trt)
 
```

### PCA Plots

TO DO

-   Re-read fundamentals of PCA

```{r}

#Plot PCA scores plot overlay
dsnv <- spec_trt

pcs <- prcomp(dsnv[,-1])

pcss <- pcs$x[,1:10]

pcss[1:6,]

plot(pcss)

#FUSI EDIT: The original code was calib <- dim(master)[1]; but this refers to the master which as defined above is based on data used to develop the model
# I changed it to refer to nwraw0 here because that is the data being used to develop the model
calib <- dim(nwraw0)[1]


#calib <- dim(master)[1]

points(pcss[1:calib,1:2], col = "red")

points(pcss[-c(1:calib),1:2], col = "blue")

var <- round(summary(pcs)$importance[2,] * 100, 1)

scores <- cbind("Calib",as.data.frame(pcs$x[,1:5])) # get first 5 principal components

names(scores) <- c("set", colnames(scores[,-1]))

scores <- as.matrix(scores)

scores[-c(1:calib),1] <- "New samples"

scores <- as.data.frame(scores)

write.csv(scores, file = "./Calib and Pred scores.csv", row.names = FALSE)

scores <- read.csv("./Calib and Pred scores.csv")

#sp <- sp +  labs(color = "set")
sp <- ggplot(scores, aes(x = PC1, y =PC2, colour = set)) +
  
  geom_point(size = 0.8, alpha = 0.85 ) +
  
  ggtitle("Calibration PCA scores plot") +
  
  
  xlab(paste0("PC1 explains ", var[1], "% total variance")) +
  
  ylab(paste0("PC2 explains ", var[2], "% total variance")) +
  
  theme_bw() +
  
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
  )
sp <- sp + theme(plot.title = element_text(hjust = 0.5))

sp <- sp + scale_color_manual(values =c("brown","orange"))

sp

ggsave(filename  = "./Soils_scores.png", height = 6, width = 6,sp)
  
```

```{r}

```

# Section 05

-   note:

    -   section 03: Kennard Stone is not used for the Cornell MIR files, because we don't need to split and determine which samples need to be sent for wet chem analysis since we already have the wet chem data (for whichever samples)

    -   section 04: Overlaying, also not used here since we don't have any sort of reference model file

## PLS Model

---
title: "05CalibrationModels_PLS_Loop_ARC"
output: html_document
date: "2023-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\_ two ways of doing predictions (1) with an existing model that was built some time ago and (2) in this case when you have the spectral and wet chem data and you have to develop the model

Reference data = wet chemistry data of the full 100 samples (being used to develop the model)

QUESTION:

-   assuming the file spectral data is reading is referring to the selected samples from section 03?

-   or which method of splitting calibration and training data is used? KS or the random setseed123 used later in this script?

```{r}
library(dplyr)
library(prospectr)
library (globals)
library(stringr)
library(ggplot2)


#FUSI EDIT
#The file being uploaded here is lilely just the cleaned spectra on the 100 biochars (since they have the reference data that is being refernced later)

# spectraldata=read.csv("/Users/vkarari/Library/CloudStorage/OneDrive-CGIAR/Documents/0All_Training/SA_ARC_Physical_Training_Aug-Sept22/ARC-MIR/ARC-MIR.csv")
# rownames(spectraldata)<-spectraldata$SSN



#FUSI EDIT: 
# since I'm combining the script into one, I don't need to reimport the data, just reference it

#Remove metadata columns when using raw file
# names(spectraldata)
# spectraldata<-spectraldata[,-2:-17]
# 
# library("stringr")
# #Ensure all the wavenumbers are rounded to one decimal place
# names(spectraldata)=round(as.numeric(str_sub(names(spectraldata),2)),1)
# names(spectraldata)[1] <- "SSN"
# 

##August 3rd Stopping point: need to work on reference data before uploading 

#Read reference data
df1=read.csv("/Users/vkarari/Library/CloudStorage/OneDrive-CGIAR/Documents/0All_Training/SA_ARC_Physical_Training_Aug-Sept22/ARC-Ref/ARC-Ref.csv")#[,c(1,8:29)]

rownames(df1)<-df1$SSN


# 
# 
# ##################################### Spectra pre-treatment ###################
# #remove the CO2 region from all data
# co2rem<-which((as.numeric(colnames(spectraldata[-1]))) < 2380 & (as.numeric(colnames(spectraldata[-1]))) > 2350)
# spectraldata.f<-spectraldata[,-co2rem]
# 
# #Mean-centering function
# center_colmeans <- function(x) {
#   xcenter = colMeans(x)
#   x - rep(xcenter, rep.int(nrow(x), ncol(x)))
# }
# 
# #SavitzkyGolay pretreatment
# spec_der<-as.data.frame(savitzkyGolay(spectraldata.f[,-1], w=17, p=2, m=1))
# 
# #Mean-centering
# spec_der.mc_trt<-center_colmeans(spec_der)
# 
# spec_trt<-bind_cols(SSN=spectraldata$SSN,spec_der.mc_trt)

###############################################################################

#FUSI EDIT: ... define what raw0 is
spectraldata<- raw0

#select only the common soil samples in spectral and wet chemie data
df.f<-df1[is.element(df1$SSN, spectraldata$SSN),]

# #Remove unwanted characters in the dataa
# df.f[sapply(df.f, grepl, pattern = "<")] <- "NA"
dnmrc <- df.f[,-1] %>% mutate_if(is.character, as.numeric)
df.f <- bind_cols(SSN = df.f[,1],dnmrc)

spectraldata.fin<-spectraldata[is.element(spectraldata$SSN, df.f$SSN),]

#Set splitting proportion for the calibration and validation data
set.seed(123)
pool=df.f[sample(nrow(df.f), round(0.3*nrow(df.f),0)), ]
pool<-pool[order(pool$SSN),]
poolid<-pool$SSN

##################### Random Forest Modelling #################################
library(randomForest)
library(caret)
library(pls)
#If package ithir is not available for your version of R
#Use the 3 below lines to install package ithir
# install.packages("devtools") 
# library(devtools)
# install_bitbucket("brendo1001/ithir/pkg") 
library(ithir)

#Available properties to predict
#Incase you want to predict only selected properties,
#get property position by running line 66. Remove hash sign between ")" and "["
#symbols on line 67. Edit properties position and run line 67.
#Always ensure position 1 in always included.
names(df1)
slprptr<-names(df1)[c(1:5,9:23)]

pred<-as.data.frame(spec_trt[,1])
colnames(pred)<-"SSN"

mdl.stats<-NULL#Model stats container

for(p in 2:length(slprptr)){
  
#Select properties to predict one at a time and remove NAs  
df.sel<-df.f %>% select(SSN, slprptr[p]) %>% na.omit

#Plot and print soil properties boxplots
boxplot(df.sel[,slprptr[p]], las=2, xlab = slprptr[p], ylab = "")
dir.create("Plots_Boxplots")
png(paste0(getwd(),"/Plots_Boxplots/",slprptr[p],".png"))
print(boxplot(df.sel[,slprptr[p]], las=2, xlab = slprptr[p], ylab = ""))
dev.off()

# #Split samples inside loop for variables with many NAs
# #Set splitting proportion for the calibration and validation data
# set.seed(123)
# pool=df.f[sample(nrow(df.sel), round(0.2*nrow(df.sel),0)), ]
# pool<-pool[order(pool$SSN),]
# poolid<-pool$SSN


#Get calibration and validation datasets
val_df<-pool
cal_df1 <-subset(df.sel, !(df.sel$SSN %in% val_df$SSN))
# threshold to exclude the extreme 5% values
cal_df <-subset(cal_df1, cal_df1[,2]>quantile(cal_df1[,2], 0.05)&cal_df1[,2] <quantile(cal_df1[,2], 0.95))
val_df1 <-subset(df.sel, (df.sel$SSN %in% val_df$SSN))
val_df <-subset(val_df1, val_df1[,2]>quantile(val_df1[,2], 0.05)&val_df1[,2] <quantile(val_df1[,2], 0.95))
val_df<-val_df[order(rownames(val_df)),]
cal_df<-cal_df[order(rownames(cal_df)),]

#Subset pre-treated spectra by available reference data
val_spec<-spec_trt[is.element(spec_trt$SSN, val_df$SSN),]
cal_spec<-spec_trt[is.element(spec_trt$SSN, cal_df$SSN),]
cal_spec<-cal_spec[order(cal_spec$SSN),]
val_spec<-val_spec[order(val_spec$SSN),]


#Get no of calibration and validation datasets
N_cal<-nrow(cal_spec)
N_val<-nrow(val_spec)

#Model data
Xcal.f=cal_spec[,-1]
Xval.f=val_spec[,-1]
dfcal.f=cal_df[,-1]
dfval.f=val_df[,-1]

###### PLSR SOC
maxc <- 25  ## number of max components
pls.md <- plsr(dfcal.f ~ ., data = Xcal.f, maxc, validation = "CV", segments = 10)#10-fold CV

## plot RMSEP vs. number of components
plot(pls.md, "val", main=slprptr[p]) 

dir.create("Components_plots")
png(paste0(getwd(),"/Components_plots/",slprptr[p],".png"))
print(plot(pls.md, "val", main=slprptr[p]))
dev.off()

## no. components to use, the one with the smallest adj RMSEP value
RMSEP.obj<-RMSEP(pls.md)
str(RMSEP.obj)

RMSEP.obj$val[1:2,1,]
nc <- as.numeric(sub("comps", "", names(which.min(RMSEP.obj$val[1,1,2:dim(RMSEP.obj$val[1:2,1,])[2]]))))
nc

#Generate relevant model name
md.nm<-paste0("pls.md.", slprptr[p], ".nc", nc)

#Rename model with the looped soil property
assign(x = md.nm, value = get("pls.md", pos = .GlobalEnv), pos = .GlobalEnv)

## predict to validation dataset
pls.prd <- predict(pls.md, ncomp = nc, newdata = Xval.f)

## Return prediction statistics
val.stats=round(goof(dfval.f, pls.prd, type = "spec"),3)
val.stats<-bind_cols(Property=paste0(Property=slprptr[p],"_val"), Comps="", N=N_val, val.stats)
val.stats

## calibration statistics
pls.pc <- predict(pls.md, ncomp = nc, newdata = Xcal.f)

pls.cal=round(goof(dfcal.f, pls.pc, type = "spec"),3)
cal.stats<-bind_cols(Property=paste0(Property=slprptr[p],"_cal"), Comps=as.character(nc), N=N_cal, pls.cal)
cal.stats

################### Get model statistics #########################
mdstats<-bind_rows(cal.stats, val.stats)

#Create model stats labels for the plot
slct.stats<-as.data.frame(t(mdstats[,c("Property","Comps","N","R2","RMSE","bias","RPIQ" )]))
names(slct.stats)<-NULL
slct.stats<-bind_cols(rownames(slct.stats),slct.stats[,2])
valbls<-paste0(c("N","R2","RMSE","bias","RPIQ"), "\n")
valsts<-paste0(c(slct.stats[2,2],slct.stats[3,2],slct.stats[4,2],slct.stats[5,2],slct.stats[6,2]))
valstats<-paste(valbls,valsts)

#Bind all looped properties model stats
mdl.stats<-bind_rows(mdl.stats,mdstats)


lgth<-length(sort(dfval.f,decreasing=F))

seq.int(sort(dfval.f,decreasing=F)[1], sort(dfval.f,decreasing=F)[lgth],length.out=4)

#Plot validation plot
plot(dfval.f,pls.prd,pch=10,
     xlab=paste('Measured',names(val_df)[2],sep="_"),
     ylab=paste('Predicted',names(val_df)[2],sep="_"), 
     xlim = range(c(dfval.f,pls.prd)),
     ylim = range(c(dfval.f,pls.prd)),
     mtext(valstats[-1],side=3, at=c(seq.int(sort(dfval.f,decreasing=F)[1], sort(dfval.f,decreasing=F)[lgth],length.out=4)))
     )   ## plot the predicted vs. measured in the validation
abline(a = 0, b = 1)


dir.create("Plots_Validationplots")
png(paste0(getwd(),"/Plots_Validationplots/",slprptr[p],".png"))
print(plot(dfval.f,pls.prd,pch=10,
           xlab=paste('Measured',names(val_df)[2],sep="_"),
           ylab=paste('Predicted',names(val_df)[2],sep="_"), 
           xlim = range(c(dfval.f,pls.prd)),
           ylim = range(c(dfval.f,pls.prd)),
           mtext(valstats[-1],side=3, at=c(seq.int(sort(dfval.f,decreasing=F)[1], sort(dfval.f,decreasing=F)[lgth],length.out=4)))
           ))
abline(a = 0, b = 1)
dev.off()


################### Predict all samples #########################
prd.smpls <- predict(pls.md, spec_trt[,-1])

prd<-as.data.frame(prd.smpls)
df.prd<-bind_cols(SSN=rownames(prd),prd[,nc])
colnames(df.prd)<-c("SSN",slprptr[p])

pred<-merge(pred, df.prd, by="SSN", all.x = T)
}

#Remove the least reliably predicted texture data (Clay,Sand or Silt)
#Recalculate the removed texture data to make Clay+Sand+Silt=100% content

#Write model statistics and predicted values to the local drive
write.csv(mdl.stats, paste0(getwd(),"/Model_Statistics.csv"),row.names = F)
write.csv(pred, paste0(getwd(),"/Predicted_Soil_Properties.csv"),row.names = F)
getwd()



############# Predicting New Spectra Using Existing Models #####################

new_spectra<-read.csv("/Users/vkarari/Library/CloudStorage/OneDrive-CGIAR/Documents/0All_Training/SA_ARC_Physical_Training_Aug-Sept22/ARC-MIR/Argd_Raw_spectra.csv")

wavenumbers <- as.numeric(str_sub(names(new_spectra),2))
names(new_spectra) <- wavenumbers
names(new_spectra)[1] <- "SSN"
rownames(new_spectra)<-new_spectra$SSN


#Pretreat new data as the calibration data was pretreated
co2rem.n<-which((as.numeric(colnames(new_spectra[-1]))) < 2380 & (as.numeric(colnames(new_spectra[-1]))) > 2350)
newspectra.fin<-new_spectra [,-co2rem.n]

spec_vnew.p<-as.data.frame(savitzkyGolay(newspectra.fin[,-1], w=17, p=2, m=1))
spec_vnew.pm<-center_colmeans(spec_vnew.p)


##### Predict new dataset using PLS model: Use the model name suffix as ncomp #####
Clay_pNEW_pls <- as.data.frame(predict(pls.md.Clay.nc11, ncomp = 11, newdata = spec_vnew.pm))

SOC_pNEW_pls <- as.data.frame(predict(pls.md.Carbon.nc14, ncomp = 14, newdata = spec_vnew.pm))

CEC_pNEW_pls <- as.data.frame(predict(pls.md.CEC.nc23, ncomp = 23, newdata = spec_vnew.pm))

PH_H2O_pNEW_pls <- as.data.frame(predict(pls.md.pH_H2O.nc14, ncomp = 14, newdata = spec_vnew.pm))

##################### Predicted dataset using PLSR model
predNEW<-cbind(new_spectra$SSN, Clay_pNEW_pls, SOC_pNEW_pls, CEC_pNEW_pls, PH_H2O_pNEW_pls)

colnames(predNEW) <- c("SSN", "Clay", "SOC", "CEC", "PH_H2O")

write.csv(predNEW,"predicted_pls_all.csv", row.names = F)


```
