## Installing Packages

```{r}

#if (!require("remotes")) install.packages("remotes")
#remotes::install_github("philipp-baumann/simplerspec")

#Info on package: 
  #https://github.com/philipp-baumann/simplerspec/ & 
  #https://github.com/philipp-baumann/simplerspec-read-filter-transform/blob/master/README.Rmd 

#install.packages(c("simplerspec","ggfortify"))
library(devtools)
#install_github("vqv/ggbiplot")
#library(simplerspec)
# Simplerspec is a universal file reader that allows one to read selected parameters instrument, optic and acquisition parameters)

library(readxl)
library(plyr)
library(dplyr)
library(tidyr)
library(ggfortify)
library(tibble)
library(here)
library(ggbiplot)
```

```{r}
files <- list.files("./../../FTIR_Spectra/Oxidized_Isotherm_Spectra/2023_03_13", full.names = TRUE)
str(files)
```

The object `files` has the data structure *atomic vector*. *Atomic vectors* have six possible basic (*atomic*) vector types. These are *logical*, *integer*, *real*, *complex*, *string* (or *character*) and *mir*. Vector types can be returned by the R base function `typeof(x)`, which returns the type or internal storage mode an object `x`. For the `files` object it is

```{r}
# Check type of files object
typeof(files)
```

## Gathering Spectra

-   Took the average of the three scans

-   Created a dataframe (df2) that shows the wavenumber and the average (of 3) absorbances for all samples

```{r}

the_names<- data.frame(list.files(path="./../../FTIR_Spectra/Oxidized_Isotherm_Spectra/2023_03_13", pattern="*.CSV", full.names=TRUE)) 

#naming the column containing file names "ID"
colnames(the_names)=("ID")

#Removing excess characters around the file names to only retain the file name and not the path
#based on the number of strings in the file name (80 in this instance)
the_names<-the_names%>%
  mutate(ID=substr(ID, 70,76))


absorbance <- list.files(path="./../../FTIR_Spectra/Oxidized_Isotherm_Spectra/2023_03_13", pattern="*.CSV", full.names=TRUE) %>% 
  lapply(read.csv,header=FALSE,col.names=c("wavenumber","absorbance")) %>%
  lapply("[",2) #%>%
 # reduce(merge,by = c("absorbance"))
absorbance

wavenumber <- list.files(path="./../../FTIR_Spectra/Oxidized_Isotherm_Spectra/2023_03_13", pattern="*.CSV", full.names=TRUE) %>% 
  lapply(read.csv,header=FALSE,col.names=c("wavenumber","absorbance")) %>%
  lapply("[",1) #%>%
 # reduce(merge,by = c("absorbance"))
wavenumber<-wavenumber[[1]]

absorbance<-do.call("cbind", absorbance)
colnames(absorbance)<-the_names[,1]
absorbance

df2<-cbind(wavenumber,absorbance) 
df2

#uploading the dataframe with the output variables 

#output_vars<-read_excel("IR_Dependent_Variables.xlsx")

```

## Pivoting Dataframe

-   creating wide formatted dataframe (spectra_wide) with ID column (SSN), column with feedstock name (feedstock1), and a column for each wavenumber with the corresponding absorbance as the entry

```{r}
spectra_long<-pivot_longer(df2,c(the_names[,1]))


newdf <- spectra_long %>%
  group_by(name,wavenumber) %>%
  summarise(value=mean(value,na.rm=FALSE))
newdf<-data.frame(newdf)


wavedf <- spectra_long %>%
  group_by(name) %>%
  summarise(wavenumber=mean(wavenumber,na.rm=FALSE))
wavedf<-data.frame(wavedf)
wavedf<-data.frame(wavedf[,2])
colnames(wavedf)<-("wavenumber")

newdf2<-cbind(wavedf ,newdf)


spectra_wide<-pivot_wider(newdf,names_from="wavenumber",values_from="value")
```

```{r}


the_names<- data.frame(list.files(path="FTIR_Spectra/2023_03_13", pattern="_1.CSV", full.names=TRUE)) 

colnames(the_names)=("ID")

the_names<-the_names%>%
  mutate(ID=substr(ID, 36,40))

feedstock<-the_names%>%
  mutate(feed=substr(ID, 1,2))
feedstock1<-as.factor(feedstock[,2])


spectra_wide<-cbind(spectra_wide[,1], feedstock1,spectra_wide[,c(2:length(spectra_wide))])
colnames(spectra_wide)[1]<-"SSN"


```

## Preliminary PCA

```{r}

#delete_out<-output_vars[c(1:6),]
#delete_in<-data.frame(spectra_wide)

#delete_RHS_pca <-prcomp (delete_in[,c(2:length(delete_in))],
                 #center=TRUE)

tot_pca <- prcomp(spectra_wide[,c(3:length(spectra_wide))],
                 center=TRUE)

summary(tot_pca)
```

```{r}
g <- ggbiplot::ggbiplot(tot_pca, #need to specify which library the command is coming from, gives errors otherwise
              obs.scale = 1,
              var.scale = 1 ,
              groups = spectra_wide$feedstock1,
              ellipse = TRUE,
              circle = TRUE,
              ellipse.prob = 0.68,
              var.axes=FALSE)
g
```

```{r}


#tot_pca_plot<-autoplot(tot_pca) #, data=spectra_wide, colour="feedstock1")

#tot_pca_plot
```

```{r}
g2 <- ggbiplot::ggbiplot(tot_pca,
              obs.scale = 1,
              var.scale = 1 ,
              groups = spectra_wide$feedstock1,
              ellipse = FALSE,
              circle = TRUE,
              ellipse.prob = 0.68,
              var.axes=FALSE)+
  geom_text(aes(label = output_vars$O_C))
g2
```

```{r}
library(downloader)
```

### Andrew Sila Script

The following is taken from : [link](https://raw.githubusercontent.com/asila/RF_PLS_Ensemble_methods/master/RF_PLS_optimal.R)

```{r}

mir<-spectra_wide[,c(1,3:length(spectra_wide))]
ref<-output_vars
m<-round(0.1*nrow(ref))
test<-sample(1:nrow(ref),m)
hout<-ref[test,]

wd<-here()

```

Trans function

```{r}

#' Derives first derivative from ab matrix
#'
#' @author Andrew Sila \email{asila@cgiar.org}

trans <- function (mir, tr = "derivative", order = 1, gap = 21, plot.spectrogram=FALSE){
     if (class(as.numeric(colnames(mir))) != "numeric") {
         stop("Invalid argument: the colnames of 'mir', which should be the waveband positions, are not coercible to class 'numeric'.")
     }
    if (as.numeric(colnames(mir)[1]) > as.numeric(colnames(mir)[2])) {
        test <- mir
        for (i in 1:nrow(mir)) {
            test[i, ] <- rev(test[i, ])
        }
        colnames(test) <- rev(colnames(test))
        mir <- test
        rm(test)
    }
    if (class(mir[1]) == "data.frame") { #Fusi update [1]
        mir <- as.matrix(mir)
    }
    #if (class(mir[1]) != "matrix") { #Fusi update [1]
    #    stop("Invalid argument: 'mir' must be of class 'matrix' or 'data.frame'.")
    }
    if (is.na(match(tr, c("derivative", "continuum removed", 
        "wt")))) {
        stop("Invalid argument: 'tr' must be 'derivative','continuum removed' or 'wt'")
    }
    if (tr == "derivative") {
        order <- round(order)
        if (is.na(match(order, c(0:3)))) {
            stop("Invalid argument: 'order' has to be an integer between 0 and 3.")
        }
        gap <- round(gap)
        if (is.na(match(gap, c(1:30)))) {
            stop("Invalid argument: 'gap' has be an integer between 1 and 30.")
        }
    }
    if (tr == "derivative") {
        transdf <- matrix(nrow = nrow(mir), ncol = ncol(mir), dimnames = list(rownames(mir), 
            colnames(mir)))
        waveb <- as.numeric(colnames(mir))
        for (i in 1:nrow(mir)) {
            transdf[i, ] <- locpoly(waveb, mir[i, ], drv = order, 
                bandwidth = gap, gridsize = ncol(mir))[[2]]
        }
    }
    if (tr == "continuum removed") {
        transdf <- matrix(nrow = nrow(mir), ncol = ncol(mir), dimnames = list(rownames(mir), 
            colnames(mir)))
        waveb <- as.numeric(colnames(mir))
        test <- mir
        for (i in 1:nrow(mir)) {
            test.1 <- cbind(waveb, test[i, ])
            test.1 <- sortedXyData(test.1[, 1], test.1[, 2])
            ch <- chull(test.1)
            ch.1 <- ch
            ch <- ch[1:(which(ch == 1))]
            ch <- sort(ch)
            ch <- c(ch, ncol(mir))
            appr.ch <- approx(test.1[ch, ], xout = test.1[, 1], 
                method = "linear", ties = "mean")
            cr <- test.1[[2]] - appr.ch[[2]]
            transdf[i, ] <- cr
        }
        transdf <- transdf[, 2:(ncol(mir) - 2)]
    }
    if (tr == "wt") {
        waveb <- as.numeric(colnames(mir))
        waveb.1024.up <- round(max(waveb))
        waveb.1024.down <- round(min(waveb))
        waveb.1024.n <- 1023
        waveb.1024.step <- (waveb.1024.up - waveb.1024.down)/waveb.1024.n
        waveb.1024 <- c()
        waveb.1024[1] <- waveb.1024.down
        for (i in 2:1024) {
            waveb.1024[i] <- round(waveb.1024.down + (i - 1) * 
                waveb.1024.step, 5)
        }
        mir.comp <- matrix(nrow = nrow(mir), ncol = length(waveb.1024), 
            dimnames = list(rownames(mir), waveb.1024))
        for (i in 1:nrow(mir)) {
            mir.comp[i, ] <- round(spline(waveb, mir[i, ], method = "natural", 
                xout = waveb.1024)[[2]], 6)
        }
        lev <- 7
        slo <- 3
        filte = "haar"
        transdf <- matrix(nrow = nrow(mir.comp), ncol = 2^lev, 
            dimnames = list(rownames(mir.comp), paste("WC_", 
                c(1:2^lev), sep = "")))
        for (i in 1:nrow(transdf)) {
            blub <- dwt(mir.comp[i, ], filter = filte)
            transdf[i, ] <- slot(blub, "W")[[slo]]
        }
    }
  waveb <- as.numeric(colnames(mir))
  if(plot.spectrogram==TRUE){
    dev.new(width = 10, height = 7)
    par(mfrow = c(2, 1))
    plot(mir[1, ] ~ waveb, type = "l", ylim = c(min(mir), max(mir)), 
        xlab = "Wavebands", ylab = "Absorption or Reflection", 
        main = "mir spectra")
    for (i in 2:nrow(mir)) {
        lines(mir[i, ] ~ waveb)
    }
  }
    if (tr != "wt") {
        waveb <- as.numeric(colnames(transdf))
        xl = "Wavebands"
        yl = "Absorption or Reflection"
    }
    if (tr == "wt") {
        waveb <- c(1:128)
        xl = "Wavelet coefficients from level 3"
        yl <- "Value wavelet coefficient"
    }
    if (tr == "derivative") {
        te <- "Derivative spectra"
    }
    if (tr == "continuum removed") {
        te <- "Continuum removed spectra"
    }
    if (tr == "wt") {
        te <- "Wavelet transformed spectra"
    }
  if(plot.spectrogram==TRUE){
    plot(transdf[1, ] ~ waveb, type = "l", ylim = c(min(transdf), 
        max(transdf)), xlab = xl, ylab = yl, main = te)
    for (i in 2:nrow(mir)) {
        lines(transdf[i, ] ~ waveb)
    }
  }
    output <- list(mir = mir, trans = transdf, transformation = tr)
    class(output) <- "trans"
    return(output)


```

RF PLS Ensemble Method

```{r}

calibrate <- function(wd,infrared.data,reference.data,hout,method = c("RF","PLS"),process = c("none","derivative", "msc", "snv","detrend")){


  # PLS Regression method
  # ----------------------------------------------------------------
  if(method ==  "PLS"){
  	
  	setwd(wd)
  	
  	mir <- infrared.data
  	
  	ref <- reference.data
  	
  	  # set preprocessing methods.
    # set preprocessing methods.
  
  if(process == "none"){
  	
  	#0. Raw
	raw <- mir
		
	colnames(raw)<-c("SSN",colnames(raw[,-1]))
	
	write.table(raw,file= paste0( process, " processed spectra.csv"),sep=",",row.names=FALSE) # same as what was read, creates a duplicate!
	}

  	
  	  	#5. First derivative
  	
  	mir1 <- as.matrix(mir[,-1])
  
  	wave <- as.numeric(substr(colnames(mir1),2,19))
  
  	prefx <- substr(colnames(mir1),1,1)[900]
  	
  	colnames(mir1) <- wave
  
  	if(process == "derivative"){
  
  	de1 <- trans(mir1,tr = "derivative",order = 1,gap = 23)
  
  	der1 <- rev(as.data.frame(de1$trans))
  
  	colnames(der1) <- paste0(prefx,wave)
  
  # Save derivative spectra.
  der1.ssn <- as.data.frame(cbind(as.vector(mir[,1]),der1))
  
  colnames(der1.ssn) <- c("SSN",colnames(der1))
  
  write.table(der1.ssn,file = paste0( process, " processed spectra.csv"),sep = ",",row.names = FALSE)
  	} 
  	
  	 # Use preprocessed table
  
  der1.ssn<-as.data.frame(read_csv(paste0( process, " processed spectra.csv")))
  
  # Merge with first derivative preprocessed spectra.
  
  	ref.mir <- merge(ref,der1.ssn,by.x = "SSN",by.y = "SSN")
	
  	rc <- colnames(ref)
  	
  	#which columns contains reference data?
  	
  	ref<-ref.mir[,rc]
  	
  	#Extract spectral predictors
  	
  	mirp<-colnames(der1.ssn)[-1]
  	
  	spectra<-ref.mir[,mirp]
  	
  	#Create two new subfolders within the current working using:
  	
  	b<-getwd()
  	
  	if(!file.exists("Models")){dir.create("Models")}
  	
  	if(!file.exists("calibration_plots")){dir.create("calibration_plots")}
  	
  	# Fit calibration models for the training set and
  	
  	# use the testing set to validate the models
  	  	
  	set.seed(67523)
  	
  	testing <- which(ref.mir$SSN%in%hout$SSN) #with hout
  	
  	#Use Kennard_Stone.
  	
  	# This is an optional step just to show distribution of spectra in a PCA space.
  	
  	sel <- kenStone(spectra,k = round(0.33*nrow(spectra)),pc = .99)
  	
  	
  	# To view selected samples, remove "#" below two lines to plot
  	
  	# plot(sel$pc[,1:2],xlab = 'PC1',ylab = 'PC2')
  	
  	# points(sel$pc[sel$model,1:2],pch = 19,col = 2)
  	
  	# points selected for calibration
  	
  	#Loop for calibration of all soil properties in the reference set starts here
  	
	msummary <- NULL
	
	hd <- colnames(ref)[-1]
	
	for (q in 1:length(hd)){
		
		refq <- which(colnames(ref)%in%hd[q])
		
		ref.q <- ref[,refq]
		
		pms.a <- NULL
		
		pred.all <- NULL
		
		cal <- cbind(as.vector(ref.q),spectra)[-testing,]
		
		val <- cbind(as.vector(ref.q),spectra)[testing,]
		
		colnames(cal) <- c(colnames(ref)[refq],colnames(spectra))
		
		colnames(val) <- colnames(cal)
		
		cal <- na.omit(cal)
		
		val <- na.omit(val)
		
		trainX <- cal[, -1]
		
		set.seed(100)
		
		colnames(cal) <- c("trainY", colnames(trainX))
		
		cal[,"trainY"] <- log(cal[,"trainY"])
	
		indx <- createFolds(cal[,"trainY"], returnTrain = TRUE)
		
		ctrl <- trainControl(method = "cv", index = indx)
		
		rf.m <- train(trainY~., method = "pls", data = cal,trControl =
		
		ctrl,tuneGrid = expand.grid(ncomp = 1:10),metric = "RMSE",preProc = 
		
		c("center", "scale"))
		
		# Get final model to compute coefficient for variation explained
		
		predi <- exp(predict(rf.m,rf.m$trainingData))
		
		y <- exp(cal[,"trainY"])
		
		#computes RMSE and R-squared values for the calibration set

		training.parameters <- round(postResample(predi,y),3)
		
		RSQ <- training.parameters[2]
		
		RMSE <- training.parameters[1]
		
		# Predict qth soil property of the holdoutset using
		
		# the MIR data and compare with the actual measurement
		
		predi.test <- exp(predict(rf.m,val[,-1]))
		
		y.test <- val[,1]
		
		#Get PCs used
		
		PCs <- rf.m$finalModel$ncomp
		
		#computes RMSE and R-squared values for the validation set

		testing.parameters <- round(postResample(predi.test,y.test),2)
		
		RSP <- testing.parameters[2]
		
		RMSEP <- testing.parameters[1]
		
		model.summary <- c(hd[q],PCs,training.parameters,testing.parameters)
		
		msummary <- rbind(msummary,model.summary)
		
		saveRDS(rf.m,file = paste0(b,"/","models/",hd[q],".rds"))
		
		pm <- as.data.frame(cbind(y,predi))
		
		colnames(pm) <- c("measured","predicted")
		
		# Create scatter plot for the predicted versus the measured - training data set
		
		p <- ggplot(pm, aes(x = measured,y = predicted)) + 
		
		geom_point(col = "black",size = 2,alpha = 0.3) + 
		
		ggtitle(paste0("Calibration for ",hd[q])) + 
		
		xlab("Measured") + 
		
		ylab("Predicted")
		
		p <- p + stat_smooth(method = lm, se = FALSE, color = 'black',alpha = 0.15)
		
		p <- p + theme(plot.title = element_text(lineheight = 3, face = "bold",
		
		color = "black", size = 20))
		
		 # this will change all text size 
		
		p <- p + theme(text = element_text(size = 20))
		
		p <- p + annotate('text', label = paste('R^2 == ',RSQ),
		
		parse = TRUE,Inf, -Inf,hjust = 2.5, vjust = -7.8)  + 
		
		annotate('text', label = paste('RMSE == ',RMSE), 
		
		parse = TRUE,Inf, -Inf,hjust = 1.8, vjust = -6.4)
		
		# Centre title
		
      	p <- p + theme(plot.title = element_text(hjust  = 0.5))
      	
      	p <- p + xlim(range(pm)) + ylim(range(pm))
      
		#Validation data
		
		pmp <- as.data.frame(cbind(y.test,predi.test))
		
		colnames(pmp)<-c("measured.test","predicted.test")
		
		# Create scatter plot for the predicted versus the measured
		
		# the validation set
		
		p2 <- ggplot(pmp, aes(x = measured.test,y = predicted.test)) + 
		
		geom_point(col = "brown",size = 2,alpha = 0.3) + 
		
		ggtitle(paste0("Validation for ",hd[q])) + 
		
		xlab("Measured") + 
		
		ylab("Predicted")
		
		p2 <- p2 + stat_smooth(method = lm, se = FALSE, color = 'brown',
		
		alpha = 0.15)
		
		p2 <- p2 + theme(plot.title = element_text(lineheight = 3,
		
		face = "bold", color = "black", size = 20))
		
		# this will change all text size 
		
		p2 <- p2 + theme(text = element_text(size = 20))
	
		p2 <- p2 + annotate('text', label = paste('R^2 == ',RSP),
		
		parse = TRUE,Inf, -Inf,hjust = 2.5, vjust = -7.8)  +
		
		annotate('text', label = paste('RMSE == ',RMSEP),
		
		parse = TRUE,Inf, -Inf,hjust = 1.8, vjust = -6.4)
		
		# Centre title
      	p2 <- p2 + theme(plot.title = element_text(hjust  = 0.5))
      	
      	p2 <- p2 + xlim(range(pmp)) + ylim(range(pmp))

		# Save calibration and validation plots
		png(file = paste0(b,"/Calibration_plots/",hd[q],".png"),
		
		height = 400,width = 800)
		
		grid.arrange(p,p2,nrow = 1)
		
		dev.off()
		
	}
	
	#colnames(msummary) <- c("Soil_properties","PCs","LOOCV_RMSEC",
      
      #"LOOCV_Rsquared", "Holdout_RMSEP","Holdout_Rsquared")
      
      write.table(msummary,file = "Model_Summary.csv",sep = ",",row.names = FALSE)
      
      # All Samples
      
      b<-getwd()
      
      if(!file.exists("Full_Models")){dir.create("Full_Models")}
      
      if(!file.exists("Full_calibration_plots")){dir.create("Full_calibration_plots")}
      
      # Begin calibration 
      
      msummary<-NULL
      
      hd<-colnames(ref[,-1])#Exclude SSN 
      
      all.predicted<-NULL
      
      for (q in 1:length(hd)) {
      	
      	refq<-which(colnames(ref)%in%hd[q])
      	
      	ref.q<-ref[,refq]
      	
      	cal<-cbind(as.vector(ref.q),spectra)
      	
      	cal<-na.omit(cal)
      	
      	trainX <-cal[, -1]
      	
      	colnames (cal) <- c("trainY",colnames(trainX))
      	
      	cal[,"trainY"] <-log(cal[,"trainY"])
      	
      	#colnames(cal)<-c(colnames(ref)[refq],colnames(spectra))
      	
      	p<-which(is.na(der1.ssn[,1]) == TRUE)
      	
      	ifelse(length(p)>0,ssn<-der1.ssn[-p,1],ssn <- der1.ssn[,1])
      	
      	ifelse(length(p)>0,der1.ssn<-der1.ssn[-p,],der1.ssn<-der1.ssn)
      	
      	#Select training and testing sets
      	
      	set.seed(100)
      	
      	indx <- createFolds(cal[,"trainY"], returnTrain = TRUE)
      	
      	ctrl <- trainControl(method = "cv", index = indx)
      	
      	rf.m <- train(trainY~., method = "pls", data = cal,
      	
      	trControl = ctrl,tuneGrid = expand.grid(ncomp = 1:10),
      	
      	metric = "RMSE",preProc = c("center", "scale"))
      	
      	#Save the model
      	
      	saveRDS(rf.m,file = paste0(b,"/","Full_Models/",hd[q],".rds"))
      	
      	#Get final model to compute coefficient for variation explained
      	
      	predi <- exp(predict(rf.m,rf.m$trainingData))
      	
      	y <- exp(cal[,1])
      	
      	#Get PCs used
      	
      	PCs <- rf.m$finalModel$ncomp
      	
      	training.parameters <- c(hd[q],PCs,round(postResample(predi,y),3))
      	
      	RSQ <- round(as.numeric(training.parameters[4]),3)
      	
      	RMSE <- round(as.numeric(training.parameters[3]),3)
      	
      	msummary <- rbind(msummary,training.parameters)
      	
      	#Training
      	
      	pm <- as.data.frame(cbind(y,predi))
      	
      	colnames(pm) <-c ("measured","predicted")
      	
      	png(file = paste0(b,"/","Full_calibration_plots/",hd[q],".png"),
      	
      	height =  600,width = 600)
      	
      	p1 <- ggplot(pm, aes(x = measured,y = predicted)) + 
      	
      	geom_point(col = "brown",size = 2,alpha = 0.3) + 
      	
      	ggtitle(paste0("Calibration for ",hd[q])) + 
      	
      	xlab("Measured") + 
      	
      	ylab("Predicted")
      	
      	p1 <- p1 + stat_smooth(method = lm, se = FALSE, color = 'brown',
      	
      	alpha = 0.15) + 
      	
      	theme(plot.title = element_text(lineheight = 3, 
      	
      	face = "bold", color = "black", size = 20)) + 
      	
      	# this will change all text size 

      	theme(text = element_text(size = 20)) +
      	
      	annotate('text', label = paste('R^2 == ',RSQ),
      	
      	parse = TRUE,Inf, -Inf,hjust = 2.5, vjust = -7.8,size = 5) +
      	
      	annotate('text',label = paste('RMSE == ',RMSE), 
      	
      	parse = TRUE,Inf, -Inf,hjust = 2.0, vjust = -6.8,size = 5) + 
      	
      	annotate('text', label = paste('PCs == ',PCs), 
      	
      	parse = TRUE,Inf, -Inf,hjust = 2.5, vjust = -3.9,size = 5)
      	
      	# Centre title
      	
      	p1 <- p1 + theme(plot.title = element_text(hjust  = 0.5))
      	
      	# Create scatter plot for the predicted versus the measured 
      	
      	# the combined dataset
      	
      	p1 <- p1 + xlim(range(pm)) + ylim(range(pm))

      	ggsave(file = paste0(b,"/","Full_calibration_plots/",hd[q],".png"),
      	
      	height = 6, width = 6, units = "in", p1)
      	
      	prediction.f <- round(exp(predict(rf.m,der1.ssn[,-1])),2)
      	
      	all.predicted <- cbind(all.predicted,prediction.f)
      	
      	}
      	
      	#Combine the predicted values together
      	
      	all.predicted.SSN <- cbind(as.vector(ssn),all.predicted)
      	
      	colnames(all.predicted.SSN) <- c("SSN",hd)
      	
      	#colnames(msummary)<-c("Soil_properties","PCs","RMSEC","Rsquared")
      	
      	#Save full model summaries
      	
      	write.table(msummary, file = "Full models summary.csv",sep = ",",
      	
      	row.names = FALSE)
      	
      	#Save the linked file
      	
      	write.table(all.predicted.SSN, file = "All predictions.csv",
      	
      	sep = ",", row.names = FALSE)
     }
   }

```

```{r}

mir<-data.frame(spectra_wide[,c(1,3:length(spectra_wide))])

ref<-output_vars
m<-round(0.1*nrow(ref))
test<-sample(1:nrow(ref),m)
hout<-ref[test,]

wd<-here()

#---


  	
#---

calibrate(wd,mir,ref,hout, method="PLS", process="derivative")
```

```{r}

```
